---
title: "Data Skills 04 - Design Features with ggplot2"
author: "Christopher Prener, Ph.D."
date: '(`r format(Sys.time(), "%B %d, %Y")`)'
output:
  github_document: default
  html_notebook: default
---

## Dependencies
This notebook requires two packages from the `tidyverse` as well as two additional packages:

```{r load-packages}
# tidyverse packages
library(dplyr)     # data wrangling
library(ggplot2)   # read and write csv files
library(ggthemes)  # additional ggplot2 themes

# manage file paths
library(here)      # manage file paths
```

## Load Data
For today's session, we'll use the `penguins` data in the `palmerpenguins` package.

```{r}
penguins <- palmerpenguins::penguins
```

## Reviewing Our Last Plot
This is our code from last week. We'll start by reviewing some of the basic elements of `ggplot2`!

```{r}
p <- ggplot(data = penguins, mapping = aes(x = bill_length_mm, y = bill_depth_mm,
                                      color = species)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ x) +
  scale_x_continuous(limits = c(30, 60), breaks = seq(30, 60, by = 5)) +
  scale_y_continuous(limits = c(12, 22), breaks = seq(12, 22, by = 2)) + 
  scale_color_brewer(palette = "Dark2", name = "Species") +
  labs(
    title = "Penguin Bill Size",
    subtitle = "Multiple Islands",
    x = "Bill Length (mm)",
    y = "Bill Depth (mm)",
    caption = "Plot by Chris Prener"
  )

p
```

Now that we've finalized our plot, we can save it using `ggsave()`:

```{r}
ggsave(filename = here("results", "bill_length_depth.png"), plot = p)
```

## Adding More Design Elements
Instead of re-typing our code over and over again today, we'll use the following code to iterate on our plot. We'll work to add the following:

1. change the colors using the `viridis` palette
  * [`ggplot2` documentation for `viridis`](https://ggplot2.tidyverse.org/reference/scale_viridis.html)
2. change the theme using both `ggplot2` themes and the `ggthemes` package
  * [`ggplot2` themes](https://ggplot2.tidyverse.org/reference/ggtheme.html)
  * [`ggthemes` themes](https://yutannihilation.github.io/allYourFigureAreBelongToUs/ggthemes/)
3. move the position of the legend
4. add facets

```{r}
p <- ggplot(data = penguins, mapping = aes(x = bill_length_mm, y = bill_depth_mm,
                                      color = species)) +
  geom_point() +
  geom_smooth(method = "lm", formula = y ~ x) +
  scale_x_continuous(limits = c(30, 60), breaks = seq(30, 60, by = 5)) +
  scale_y_continuous(limits = c(12, 22), breaks = seq(12, 22, by = 2)) + 
  scale_color_brewer(palette = "Dark2", name = "Species") +
  labs(
    title = "Penguin Bill Size",
    subtitle = "Multiple Islands",
    x = "Bill Length (mm)",
    y = "Bill Depth (mm)",
    caption = "Plot by Chris Prener"
  )

p
```

Now that we've finalized our plot, we can save it using `ggsave()`:

```{r}
ggsave(filename = here("results", "bill_length_depth_new.png"), plot = p)
```

## Experimentation
Now, you try adjusting the plot above to fit new data, comparing `flipper_length_mm` and `body_mass_g`. Save your plot when you're done!

```{r}
p <- ggplot(data = penguins, mapping = aes(x = flipper_length_mm, y = body_mass_g,
                                      color = species)) +
  geom_point() +
  scale_color_brewer(palette = "Set1", name = "Species") +
  scale_x_continuous(
    limits = c(170, 240), 
    breaks = seq(170, 240, by = 10)
  ) +
  scale_y_continuous(
    limits = c(2500, 6500), 
    breaks = seq(2500, 6500, by = 500)
  ) +
  labs(
    title = "Body Mass and Flipper Length by Species",
    x = "Flipper Length (mm)",
    y = "Body Mass (g)"
  )

p
```

Now that we've finalized our plot, we can save it using `ggsave()`:

```{r}
ggsave(filename = here("results", "flipper_mass.png"), plot = p)
```
